A Pandemic-Scale Ancestral Recombination Graph for SARS-CoV-2

### 2025-10-16 latex/main.tex
Update for new 3SEQ sc2ts results

We validated the recombination events in the sc2ts ARG by running 
3SEQ~\cite{Lam2018improved} on the putative recombinant 
sequence trios (STAR Methods), 
and 383 of the 855 (45\%) were classified as recombinant.
Validation status correlates strongly with our QC filter, with 
88\% of the 354 QC-passing recombinants 
and 14\% of the QC-failing recombinants
being classified as recombinant.
The 12\% of QC-passing recombinants
not validated by 3SEQ are also among the most marginally supported
(see next section; Figure~\ref{fig:recombinant_evidence}).
In addition to 3SEQ, we ran the CovRecomb~\cite{Li2024CovRecomb} 
and rebar recombination detection methods on the putative recombinant 
sequences, summarized in Table~\ref{tab:recombinant_validation}.
CovRecomb has a substantially lower validation rate at 25\% 
among the QC-passing recombinants. This is expected, as 
CovRecomb and is specifically intended to detect interlineage recombinants
and been previously shown to be
less sensitive than other methods~\cite{Li2024CovRecomb}.
Rebar validation is intermediate, at 57\% among QC-passing recombinants.
(See also Figures~\ref{fig:recombinant_evidence_cov_recomb},
and \ref{fig:recombinant_evidence_rebar}.)
% We don't talk about RIPPLES results yet as we haven't discussed it, a
% and that would be confusing.

@@

Figure~\ref{fig:recombinant_evidence}A plots
the number of mutations required
in the recombinant versus non-recombinant
Viterbi solutions (upper scatter plot)
and the resulting distribution of the number of
averted mutations (lower histogram)
over all 354 recombination events that pass QC.
Also shown is the 3SEQ classification status of these recombinants.
All of the 12\% of recombinants that do not pass 3SEQ validation are
from the most weakly supported, with 4 or 5 averted mutations.
Figures~\ref{fig:recombinant_evidence_cov_recomb} and
~\ref{fig:recombinant_evidence_rebar} 
plot the same information for CovRecomb and rebar validation,
showing that their validation status tends to be function
of the divergence between the recombinant parents.

@@

\begin{table}
\caption{Validation of sc2ts recombinants by 3SEQ, CovRecomb and rebar over 
all 855 recombinants and the 354 passing and 501 failing QC.
}
\label{tab:recombinant_validation}
\centering
\begin{tabular}{llll}
\toprule
Method & Total & QC Pass & QC Fail \\
\midrule
3SEQ & 383 & 71 & 312 \\
CovRecomb & 92& 90 & 2\\
rebar & 213& 203 & 10 \\
\bottomrule
\end{tabular}
\end{table}

>>>>

We validated the recombination events in the sc2ts ARG by running 
3SEQ~\cite{Lam2018improved} on the putative recombinant 
sequence trios (STAR Methods), 
and 484 of the 855 (57\%) were classified as recombinant.
Validation status correlates strongly with our QC filter, with 
95\% of the QC-passing recombinants 
and 29\% of the QC-failing recombinants
being classified as recombinant.
In addition to 3SEQ, we ran the CovRecomb~\cite{Li2024CovRecomb} 
and rebar recombination detection methods on the putative recombinant 
sequences, summarized in Table~\ref{tab:recombinant_validation}.
CovRecomb has a substantially lower validation rate at 25\% 
among the QC-passing recombinants. This is expected, as 
CovRecomb and is specifically intended to detect interlineage recombinants
and been previously shown to be
less sensitive than other methods~\cite{Li2024CovRecomb}.
Rebar validation is intermediate, at 57\% among QC-passing recombinants.
(See also Figures~\ref{fig:recombinant_evidence_cov_recomb},
and \ref{fig:recombinant_evidence_rebar}.)
% We don't talk about RIPPLES results yet as we haven't discussed it, a
% and that would be confusing.

@@

Figure~\ref{fig:recombinant_evidence}A plots
the number of mutations required
in the recombinant versus non-recombinant
Viterbi solutions (upper scatter plot)
and the resulting distribution of the number of
averted mutations (lower histogram)
over all 354 recombination events that pass QC.
% The 5\% of QC-passing recombinants
% not validated by 3SEQ are also among the most marginally supported
% (see next section; Figure~\ref{fig:recombinant_evidence}).
Also shown is the 3SEQ classification status of these recombinants.
All of the 12\% of recombinants that do not pass 3SEQ validation are
from the most weakly supported, with 4 or 5 averted mutations.
Figures~\ref{fig:recombinant_evidence_cov_recomb} and
~\ref{fig:recombinant_evidence_rebar} 
plot the same information for CovRecomb and rebar validation,
showing that their validation status tends to be function
of the divergence between the recombinant parents.

@@

\begin{table}
\caption{Validation of sc2ts recombinants by 3SEQ, CovRecomb and rebar over 
all 855 recombinants and the 354 passing and 501 failing QC.
}
\label{tab:recombinant_validation}
\centering
\begin{tabular}{llll}
\toprule
Method & Total    & QC Pass & QC Fail \\
\midrule
3SEQ        & 484 & 338     & 146 \\
CovRecomb   & 92  & 90      & 2\\
rebar       & 213 & 203     & 10 \\
\bottomrule
\end{tabular}
\end{table}
